---
title: "Chicken Weights EDA"
subtitle: "A report to analyze the chicken weights dataset."
author: "Ahmed Almohammed"
output:
  html_document:
    df_print: paged
    code_folding: hide
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Part I: General understanding of the dataset

The dataset contains the following variables:

```{r}
ChickWeight
```

Here is a random sample of the data set

```{R}
set.seed(1234)
ChickWeight %>% 
    slice(sample(1:nrow(.), 10, replace = FALSE)) %>% 
    knitr::kable()
```


Before we begin, we'll remove any chicks that didn't survive until the end of the study. We'll consider them separately later on.

```{r}
# Filter the chicks
ChickWeight  %>% 
    filter(Time == max(Time)) %>% 
    pull(Chick) -> chick_list

# Make a new dataset with only the required chicks
ChickWeight %>% 
    filter(Chick %in% chick_list) -> ChickWeight_Time21
    

ChickWeight %>% 
    filter(!(Chick %in% chick_list)) -> ChickWeight_Not21

```


```{R collapse=TRUE}
# view the summary of the dataset
# summary(ChickWeight)

# Plot the dataset to view it (Point Plot)
ggplot(ChickWeight_Time21, mapping = aes(Time, weight, color = Diet)) + 
    # geom_point(shape = 16, alpha = 0.5) +
    geom_line(aes(group = Chick), alpha = 0.5) +
    facet_grid(. ~ Diet) +
    stat_summary(fun.data = mean_sdl, fun.args = list(mult = 1), color = NA, fill = "black", alpha = 0.25, geom = "ribbon") +
    stat_summary(fun.data = mean_sdl, fun.args = list(mult = 1), color = "black", geom = "line") +
    scale_x_continuous(breaks = unique(ChickWeight$Time))
```


```{R collapse=TRUE}
# Plot the dataset to view it (Box Plot)
ggplot(ChickWeight, mapping = aes(x=Diet, y=weight)) +
    stat_boxplot(geom ='errorbar') + 
    geom_boxplot() +
    NULL
```


## Part II: Deeper understanding of the dataset
With the above information, we can proceed to exploring the relations between the variables for additional insights.

We will start with grouping the dataset based on the type of `Diet` variable.
```{R collapse=TRUE}
# Now, we create a tibble that shows us the details of the chickens
# when grouped by the diet they were fed on
ChickWeight %>% group_by(Diet) %>% summarise(number_of_chicks = n_distinct(Chick))
```

We can see that there are 4 types of `Diet` that were tested on a total of 50 `Chick`, with :

* **20 chickens following diet 1**
* **10 chickens following diet 2**
* **10 chickens following diet 3**
* **10 chickens following diet 4**

Now, we group them based on the `Time` variable.

```{R collapse=TRUE}
ChickWeight  %>% group_by(Time)  %>% summarise(number_of_chicks = n_distinct(Chick))
```

From the above examination, we can see that the `Diet` was for a period of 21 days,
where at the start there were the full number of `Chick`, total of 50 `Chick`. However, towards the end, we ended up
with 45 `Chick`! We can also notice a positive correlation between time and weight, where the weight of the `chick` tends to increase
along the `time` interval.

With this new information, we see that some of the chicks didnt complete the full 21 days, and so
we need to clean the data from these 5 `Chick` that did not make it to the 21 day.

```{R collapse=TRUE}



# Check the result
ChickWeight_filtered  %>% group_by(Time)  %>% summarise(number_of_chicks = n_distinct(Chick))
```

We are now able to contiue through the process of analyzing the dataset as we cleaned it off from the unrequired data 
of the 5 `Chick` that didnt complete the specified number of days of the `Diet`

We can now plot our filtered dataset. We will start with a box plot.

```{R collapse=TRUE}
# Plot a boxplot of the data
ggplot(ChickWeight_filtered, mapping = aes(x=Diet, y=weight)) +
    stat_boxplot(geom ='errorbar') +
    geom_boxplot()

# Check the highest weight in the diet groups
ChickWeight_filtered  %>% group_by(Diet)  %>% summarise(max_weight = max(weight))
```

We notice the following:-

* The median is very low in `Diet1` compared to the other diets.
* Both `Diet3` and `Diet4` have no outliers.
* Both `Diet3` and `Diet4` have a high median value compared to `Diet1` and `Diet2`

## Part III: Conclusion

In conclusion, we can sum up the above analysis in the following main points:

> There is a positive linear relationship between `time` and `weight` of the `chick`.
> As the time passes, the chicks in all diet increase in weight.
> `Diet3` makes the `chick` fatter quickly, and `Diet1` makes the `chick` fatter slowly.